Synthesis of Linear Functional Programs

### 2021-07-13 article/article.tex
WIP: Left focus refinement types
some notes

\usepackage{cmll}
\usepackage{mathpartir}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{hyperref}

@@

\title{Synthesis of Linear Functional Programs}
\author{Rodrigo Mesquita \and Bernardo Toninho}
\date{April 2021}
\institute{NOVA School of Science and Technology}

@@

\section{Formal System}\label{sec:formal_system}

The development of \synname\ was initially found on an almost direct reading of
concrete inference rules and theoretical principles (~\cite{frank pfenning notes}); and work on its environment
was also guided by formal systems and methods (e.g. the Hindley-Milner type
inference~\cite{}).  However, with the addition of recursion and richer types to
the synth and language, formal guidelines to support the implementation become
scarcer, or more complex. Accordingly, rather than derived from explicit rules,
solutions appear first to tame the extensions' inherent complex problems (such
as infinite recursion), and implicit rules come only as a byproduct.

With this section we intend to formalize the implicit rules that guide the
synthesis of more expressive specifications, alongside the already explicit
rules that model the core of the synth process, resulting in a \emph{sound} set
of rules that hopes to serve as a source of ideas on taming synthesis of
recursive functions with richer types in a linear context, as well as describe
the system in enough detail for the synthetiser to be reproducible by a
theory-driven implementation.

@@

>>>>

\usepackage{cmll}
\usepackage{mathpartir}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{hyperref}
\usepackage{todonotes}

@@

\title{Synthesis of Linear Functional Programs}
\author{Rodrigo Mesquita - Nº XXXXX \\
    Supervisor: Bernardo Toninho}
\date{July 2021}
\institute{NOVA School of Science and Technology}

@@

\section{Synthesis with Linear Types, Polymorphism and Refinements}\label{sec:formal_system}
\todo[inline]{Encurtar titulo, mas algo nesta onda que descreve
  ``tudo''}

\todo[inline]{Aqui parece-me que deves comecar por explicar o que vais
  fazer nesta seccao. Sinto falta da sintaxe da linguagem,
  possivelmente incrementalmente. Comecava por listar os tipos/termos
  do lambda linear ``simples'', depois em cada subseccao acrescentava
  as construcoes novas e como a sintese funciona para esse
  contexto. Enfatizar que a novelty é a combinacao destas features todas.}

The development of \synname\ was based on an almost direct reading of
concrete inference rules and theoretical principles (~\cite{frank
  pfenning notes});
\todo{aqui queres dizer que começaste por te basear
  num sistema de focusing para a logica linear (era sequer linear a
  formulacao que comecaste a seguir?) O que são theoretical
  principles?}
and work on its environment was also guided by formal
systems and methods (e.g.~Hindley-Milner type
inference~\cite{}).\todo{esta frase parece-me desnecessaria. O que
  queres dizer neste paragrafo, parece-me, é que partes de um sistema
  focused da literatura que depois vais extender, inspirando te em
  tecnicas standard de proof search / inferencia.}

However, with the addition of recursion and
richer types to the synthesis and language, formal guidelines
\todo{nao sao formal guidelines, nao existem trabalhos a combinar
  estas features todas, essa parte é importante. Além de que 
ao incluir estas coisas deixamos de estar completamente em proof
search, pois muitas destas coisas (e.g. recursao) nao sao
necessariamente sound logicamente.}
  Tenta usar termos 
  especificos. Richer types? Richer como?} to support
the implementation become scarcer, or more complex \todo{nao existem
  ou sao complicados? ambos? citacao?}. Accordingly,
rather than derived from explicit rules, solutions appear first to
tame the extensions' inherent complex problems (such as infinite
recursion), and implicit rules come only as a byproduct. \todo{ Nao
  percebo o que queres dizer aqui. O que sao explicit rules e implicit
  rules? Ou há regras ou nao há. O que estas a dizer é que inventaste
  tecnicas de sintese para estes casos, que correspondem as regras que
  vais apresentar. E tao explicito ou implicito como o que esta na literatura.}

With this section we aim to formalize the implicit \todo{mais uma vez\dots} rules that guide
the synthesis of more \todo{nao se faz sintese de specs, mas sim a
  partir de specs} expressive specifications, alongside the already
explicit rules that model the core of the synth \todo{synthesis} process, resulting in
a \emph{sound} set of rules that hopes to serve as a source of ideas
\todo{ nao me parece\dots As regras caracterizam o que esta
  implementado, e esse o objectivo, nao e ser uma fonte de ideias\dots }
on taming synthesis of recursive functions with richer types in a
linear context, as well as describe the system in enough detail for
the synthetiser to be reproducible by a theory-driven implementation.

@@

\mypara{Memoization} ...
